A few stickleback suffice for the transport of alleles to new lakes

### 2019-09-18 review-responses.tex

\plr{IMPORTANT: don't reorder or delete ``points'' below - it messes up the automatic numbering!}

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
This is a very well done work addressing the fundamental problem of local
adaptation in metapopulations; stickleback are a nice study system but the
findings are relevant for a wide variety of systems. I enjoyed it very much,
and my only wish is that the simulation has been run more than once. This is
not necessary for speed of adaptation - within-simulation replication of lakes
is totally sufficient for this - but for aspects related to genome scanning,
numbers of adaptive alleles in a metapopulation, and haplotype-wise effect
sizes (table 1), since (especially at high M) the number of adaptive alleles is
very few across the whole metapopulation. That said, I do not insist on
rerunning the simulation, since it must be very computationally intensive.
\end{quote}

\begin{point}{\revref}
    \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
    as migration increases - higher rates of migration allow adaptive alleles of
    larger effect to travel more efficiently through the population, even though
    they are deleterious in the ocean.''}
    Very cool. It gives a very intuitive guideline as to which genetic
    architecture of local adaptation to expect given a certain migration rate.
    There is one neat aspect of this result that is perhaps worth better
    emphasizing (please correct me if I an : detectable Fst peaks at truly adaptive
    alleles can only be very few. If we had replicate runs we could say how many
    true Fst peaks, on average, we see at M=1 and M=10, Many published studies
    (particularly in marine animals) report dozens if not hundreds Fst outliers
    throughout the genome; your simulation here suggests that most of them have to
    be false positives of some kind, especially since migration in the sea is
    likely to be high.
\end{point}

\reply{
We agree with the reviewer that this is an important point, 
and that our results are consistant with the above hypothesis. 
However, to quantify and assert that previous empiracal studies 
are reporting false-positives would require many 
more simulations (and \textit{careful} analysis), which we believe is outside the rhelm of this paper.
}

@@

\begin{point}{}
Also, about Fig. S1: I am confused what is ``statistical power'' here :
``likelihood that a SNP will be predicted to have an effect on phenotype when
there is an effect to be detected''. In this model every SNP has at least *some*
effect on phenotype, so what defines ``effect to be detected'' then?
\end{point}

\reply{
TODO: Peter, you might be able to best describe this most clearly.
}

\begin{point}{\revref}
    \textit{``Surprisingly, for all cases with the exception of M = 0.01, we found that the
    majority of adaptive allele origins traced back the original generation of
    inhabitants in the lake (Figure 7)''}
    This is really cool and counter-intuitive,
    but maybe it is not too surprising since first-arriving (``captured'') alleles
    are simply much more numerous relative to the alleles that trickle in later,
    *before the adaptation is complete*? Indeed, initial colonization involves
    arrival of 200 individuals in a single generation, which is equivalent to 200/M
    generations of subsequent immigration. Considering that at any appreciable M
    ($>=0.1$) adaptation happens very rapidly (fig 3 and S3), there might simply be
    not enough additional immigrant alleles to make a contribution. I wonder how
    this would work if colonization of a lake was a more gradual process, say 2
    individuals per generation (not all 200 at once)? Would the early-arriving
    alleles still have advantage? (I would bet yes)
\end{point}

\reply{
We agree that this is an important point to explore, and a good observation.
We have appended to the section on migration to more clearly address this \llname{disc_colon}.  
However, the available code could certainly be used in further studies to investigate this question
more in-depth.
}

@@

\begin{point}{}
Latin binomial of the focal model species appears only in the references. It would make sense to incorporate it in the abstract and introduction given that the entire ms is focused on this species.
\end{point}

\reply{
    Good point; we've added it to the first use of the name after the abstract \revref.
}

\begin{point}{Title:}
I would make it clear that this is about the ``threespine stickleback'' - there are many different stickleback species.
\end{point}

\reply{
    We believe the first sentence in the abstract makes this clear to the reader.
}

@@

\begin{point}{\revref}
``to millions of independently derived''. Do the references you cite say so? Can you really back up your claim of millions of freshwater populations?
\end{point}

\reply{
    This order of magnitute estimate is roughly based upon the length of the coastal range and coastal lake density. Because there is no concrete evidence, we have re-worded this \llname{hundreds-thousands-or-more}. 
}

@@

\begin{point}{\revref}
``heterogenous nature of divergence''. I think you should cite here the first (paper using global sampling) showing evidence for heterogenous nature of this genetic divergence: DeFaveri et al. (2011). Global analysis of genes involved in freshwater adaptation in threespine sticklebacks (Gasterosteus aculeatus). Evolution, 65(6), 1800-1807.
\end{point}

\reply{
    Thank you for pointing this out. The citations has been added.
}

@@

\begin{point}{\revref}
P2. ``stickleback in them'' - threespine stickleback in them
\end{point}

\reply{
    Fixed.
}

\begin{point}{p.3 Methods}
From here on, the ms is generally very thinly referenced. In particular, it remains unclear to reader where the many equations come from: are they something authors have derived themselves (I guess not?), or who did that?
\end{point}

\reply{
    All equations explaining dynamics of the simulation model (outside of \llname{Fst}) were chosen explicitely by the authors as input to SLiM. We have added citations for the population genetics theoretical expectations (\llname{Gil_cite} \llname{Fst}) as well as clarified the derivation of equations in the Methods section \llname{clarify_equations}.
}

@@

\begin{point}{p.5}
One is left to wonder if it realistic to think that the marine population is as big as 25 freshwater populations. Admittedly, this is discussed to some extent in the discussion, but given the abundance of the threespine stickleback in many seas (there is actually quantitative estimates of this in literature), it seems plausible that the marine population(s) is orders of magnitudes larger than that of typical freshwater population, even if combined. I guess a more quantitative view on this in light of literature would be something worth attempting.
\end{point}

\reply{
    TODO
}

\begin{point}{\revref}
    Where does the mutation rate $10^{-10}$ come from? Why this particular value, and how are the results influenced by choice of this?
\end{point}

\reply{
    This was chosen as an arbitrary, yet realistic, rate at which you could expect incoming mutations to affect phenotype. 
At $10^{6}$ loci where these mutations can arise at a rate of $10^{-10}$, and $5000$ diploid individuals in each population, you can expect $\approx 1$ mutations to arise in each population, each generation, which impact the trait value.
Another way to look at this is that it is a per-nucleotide mutation rate of $10^{-8}$ with a total target size of $10^{4}$. 
Because we hold selective pressure (optimum trait value) constant and stack mutations, mutation rate will only affect the speed of adaptation through the use of ``de-novo'' mutation. 
While this comparison is an important point in our ms,
the authors believe this rate is, if anything, a generous assumption to make.
We have re-worded and appended to the ms to hopefully make this more clear. \llname{mu_exp}
}

@@

\begin{point}{}
P4. It is bit unclear here how LD as was dealt with in the simulations. Are all the loci assumed to be unlinked?
\end{point}

\reply{
    All chromsomes are linear with a per-nucleotide recombination rate of $10^{-8}$. This has been clarified here \revref.
}

@@

\begin{point}{\revref}
What happens if generations are overlapping, as is the case in sticklebacks?
\end{point}

\reply{
The authors expect no fundemental difference between results yeilded from WF model vs. non-WF model.
}

@@

\begin{point}{\revref}
Here the authors say that a different model in SLiM would have been more appropriate to be used, but that they ``do not expect this to influence the results''. This does not sound very good - why not to check this? Why not to rerun the analyses if a more appropriate model is available?
\end{point}

\reply{
In a perfect world, we would re-run and compare. 
However, to re-write this model as non-WF would require a relatively significant amount of change to the code as well as much more computational power to re-run the analysis.
With no fundemental differences expected between the two models other than one being slightly more realistic, the authors opted to stick with WF model. 
We agree the ms makes this slightly confusing, so we have altered the wording a bit here \llname{realistic}.
}

@@

\begin{point}{Fig 1}
The modeled scenario is probably correct over shorter geographic distances. But there is plenty of empirical evidence to show that marine threespine stickleback populations over larger geographic distances are not genetically homogenous. I think it would be appropriate to acknowledge this somewhere in manuscript.
\end{point}

\reply{
The authors agree, it should be made clear that the oceanic population is not panmictic but rather has one-dimential continuous spatial structure.
We have clarified this here \llname{continuous}.
}

@@

\begin{point}{\revref}
``The efficacy of genomic scans of selection depends on gene flow''. I found this section rather uninformative: all this has been studied in quite detail already in many theoretical (and empirical) papers. This could be shortened. And references to earlier findings would be in place here too - there is no single reference to earlier works here at all.
\end{point}

\reply{
The authors believe this finding supports existing studies, specifically, how stronger genetic drift with less migration leads to higher background $F_{ST}$.
However, for our study, showing how greater sharing of adaptive alleles provides a shared signal across populations seemed necessary for a study of selection on standing variants.
Thus, the authors have decided to leave this section as is.
TODO: Looks like we could use some citations, I was unable to find any good ones :(.
}

@@

\begin{point}{\revref}
``Simulation results align with..'' It is nice that the results are referenced back to theory. I was left to think that it might have been better to do this by combining this section with the actual results. Now the structure of the ms appears to be: 1) Results \& Discussion, 2) Discussion. Furthermore, references to theory are given in the first paragraph, but not after this. Readers do not know where all the equations are derived from.
\end{point}

\reply{
The authors thought about connected it back to the results and decided it would be more confusing to the reader. 
As above, we have decided to reference Gillespie throughout this section.
}

@@

\begin{point}{\revref}
``we found that standing..'' Is ``found'' right word here? You made an inference, even a guess, that this was so. But you did not prove this?
\end{point}

\reply{
TODO
\jgg{Justify that this is actually a finding, by adding a sentence to the bit on masking to make this more concrete.} 
}

@@

\begin{point}{\revref}
``Population size''. Here it is argued that population sizes of marine and freshwater populations are not so different. I found it hard to believe this in the light of quantitative data there is. Could it simply be so that marine population is actually very much larger, but structured (as there is genetic evidence for this in many papers) so that the fraction feeding the freshwater colonizations is to some degree isolated from ones feeding the freshwaters in a particular area? This would in effect make relevant portion of the marine population smaller, without need to postulate that the total marine population equals in size to freshwater populations, and that there are millions of lakes inhabited by three spine sticklebacks. Note also that in introduction it was claimed millions of freshwater populations, here you say ``if not millions'' - inconsistent.
\end{point}

\reply{
TODO
\jgg{This is clearly tied to the above comments about pop size, not sure how to answer these effectively}
}

### 2019-09-18 Stickleback_Paper.tex
finished reviwer comments for 1 & 2

The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
has given rise to millions of independently derived freshwater populations \revpoint{2}{3}
in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
and the much more extensive use of standing genetic variation than once thought 
\citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, stickleback in them \revpoint{2}{5}
frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
mirrored those previously found to differ between ocean and freshwater populations that
that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure.
Certain aspects of the model remain simplistic due to computational constraints;
in particular, total population sizes are much smaller than in reality,
although we may capture a good picture of local dynamics.

@@

Each individual carries two linear chromosomes, each of size $10^8$ loci. 
Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$. Subsequent mutations at the same locus replace the previous allele.

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations \revpoint{2}{10}
and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation. 
(A better implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
which was not available when we wrote the simulations; but we do not expect the difference to affect results.)
To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$. No migration occurs among the freshwater populations. All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''), and parents from the same habitat with probability $1-m$.
The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The offspring then lives in the same lake as the parents. Parents for each non-migrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
Parents for each freshwater migrant are chosen in the same way as for non-migrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent. Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

@@

We used neutral mutations to calculate $F_{ST}$ for each locus, which we then averaged in windows as a measure of between-population relative differentiation. 
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, 
then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.

@@

Figure~\ref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%. At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes. 
As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.

@@

We have thus far found that the speed of adaptation depends strongly on the degree to which alleles can be shared between populations. 
However, the \textit{origin} of the alleles underlying the phenotype is still unknown. 
What is the nature of selection on standing genetic variants? 
In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of standing genetic variation a necessity for rapid adaptation of the population?
To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (Figure \ref{fig:Origin}). 
In our simulations,
any reasonable subset of the ancestral population has the potential for rapid adaptation without the need post-colonization hybridization events.

@@

\revpoint{2}{14}
How do our results compare to what is expected from population genetics theory?
Our simulations included many loci under selection and in linkage to many other selected alleles,
which makes precise calculation impossible.
Nonetheless, rough calculations based on simple population genetics theory
-- with the benefit of hindsight --
turn out to describe qualitatively most of the aspects of adaptation we observed above.
Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche},
but doing so is beyond the scope of this paper

@@

\paragraph{Migration}
The key quantity regulating the amount of standing variation in the ocean is the \emph{downstream} migration rate, from lakes to the ocean. 
How important is the upstream migration rate?
If sufficient genetic variation is not present in a new lake initially, it must appear either by new mutation or by migration. 
Since a proportion $m$ of each lake is composed of migrants each generation, 
it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization. 
This implies a dichotomy: either 
(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
(b) migration is low, so adaptation takes many multiples of $1/m$ generations. 
Since in our model lower migration also reduces the amount of variation available in the ocean, 
we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.

>>>>

The ancestral marine form of threespine stickleback fish (\textit{Gasterosteus aculeatus}) \revpoint{2}{1}
has given rise to hundreds of thousands if not millions of independently derived freshwater populations \revpoint{2}{3} \llabel{hundreds-thousands-or-more}
in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes \revpoint{2}{4}
and the much more extensive use of standing genetic variation than once thought 
\citep{schluter2009genetics, DeFaveri_2011, DeFaveri_2013, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, threespine stickleback in them \revpoint{2}{5}
frequently evolve similar phenotypes \citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
mirrored those previously found to differ between ocean and freshwater populations that
that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details and equations of the model were motivated by current understanding of threespine stickleback evolutionary, history and demography, as well as population genetics theory \citep{Gillespie_1998}. \llabel{clarify_equations}
The following model included divergent selection in the two habitats which had substantial spatial structure.
Certain aspects of the model remain simplistic due to computational constraints;
in particular, total population sizes are much smaller than in reality,
although we may capture a good picture of local dynamics.

@@

Each individual carries two linear chromosomes, each of size $10^8$ loci,
with recombination events occurring at a intralocus rate of $10^{-8}$ every generation. \revpoint{2}{9}
%Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation \revpoint{2}{8}
%in ten regions of $10^5$ loci each, spread evenly along the chromosome. 
Mutations that can affect the trait under selection can occur at rate $10^{-8}$ with a target size of $10$ Kb, (spread evenly among $1$ Mb) across each individual genome. \revpoint{2}{8}
With 5000 (diploid) individuals per population, this equates to $\approx 1$ mutation, per population, per generation, which impact individual trait value. \llabel{mu_exp}
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$. Subsequent mutations at the same locus replace the previous allele.

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations \revpoint{2}{10}
and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation. 
(A more realistic implementation would use SLiM's ``non-Wright-Fisher'' model type, \revpoint{2}{11}
which was not available when we wrote the simulations; but we do not expect the difference to affect results.) \llabel{realistic}
To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo} the marine population is continuous (unidimensional) while lakes are discrete. \llabel{continuous}
Dispersal occurs both locally along the coastline within the marine habitat, as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$. 
No migration occurs directly between the freshwater populations. 
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''), and parents from the same habitat with probability $1-m$.
The first parent of each non-migrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The offspring then lives in the same lake as the parents. Parents for each non-migrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
Parents for each freshwater migrant are chosen in the same way as for non-migrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent. Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

@@

We used neutral mutations to calculate $F_{ST}$ for each locus, which we then averaged in windows as a measure of between-population relative differentiation. 
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, 
then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$ \citep{Gillespie_1998}. \llabel{Fst}

@@

Figure~\ref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%. At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes. 
As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean. \revpoint{1}{1}

@@

We have thus far found that the speed of adaptation depends strongly on the degree to which alleles can be shared between populations. 
However, the \textit{origin} of the alleles underlying the phenotype is still unknown. 
What is the nature of selection on standing genetic variants? 
In our simulations migration from marine individuals into the new lakes continues throughout introduction and adaptation: but is consistent influx of standing genetic variation a necessity for rapid adaptation of the population?
To answer this, we traced the genealogy of \textit{all} adaptive alleles in \textit{all} individuals from the introduced lakes, after local adaptation.
Surprisingly, for all cases with the exception of $M = 0.01$, we found that the majority of adaptive allele origins traced back the original generation of inhabitants in the lake (Figure \ref{fig:Origin}). \revpoint{1}{3}
In our simulations,
any reasonable subset of the ancestral population has the potential for rapid adaptation without the need post-colonization hybridization events.

@@

\revpoint{2}{14}
How do our results compare to what is expected from population genetics theory?
Our simulations included many loci under selection and in linkage to many other selected alleles,
which makes precise calculation impossible.
Nonetheless, rough calculations based on simple population genetics theory 
-- with the benefit of hindsight --
turn out to describe qualitatively most of the aspects of adaptation we observed above \citep{Gillespie_1998}. \llabel{Gil_cite}
Because we model stabilizing selection on an additive trait controlled by a moderately large number of loci within each population, more precise expectations might be obtained through quantitative genetics \citep{svardal2014general} or even Fisher's geometric model \citep{barton2001hybridization,chevin2014niche},
but doing so is beyond the scope of this paper

@@

\paragraph{Migration}
The key quantity regulating the amount of standing variation in the ocean is the \emph{downstream} migration rate, from lakes to the ocean. 
How important is the upstream migration rate?
If sufficient genetic variation is not present in a new lake initially, it must appear either by new mutation or by migration. 
Since a proportion $m$ of each lake is composed of migrants each generation, 
it takes $1/m$ generations until the genetic variation introduced by migrants equals the amount initially present at colonization. 
This implies a dichotomy: either 
(a) migration is high, and adaptation is possible using variants present at colonization or arriving shortly thereafter, or 
(b) migration is low, so adaptation takes many multiples of $1/m$ generations. 
Since in our model lower migration also reduces the amount of variation available in the ocean, 
we expect very little contribution of subsequent migration across any value of $m$, as seen in Figure~\ref{fig:Origin}.
It remains unknown how different colonization schemes would impact the origin of adapted alleles, 
but it stands to reason based upon our results that early arriving alleles are favored as adaptive alleles in any scenario. \llabel{disc_colon}
